{\LARGE On the Use of Generic Types for Smart Contracts}

### 2022-04-27 cluster_computing/001_abstract.tex
Further edits to the Cluster Computing version

Smart contracts are pieces of code that are deployed and executed in the context of a blockchain infrastructure in order to automatically enforce some effects when particular events occur. The writing of smart contracts is a complex and critical activity that can benefit from the use of high-level features of programming languages, like generics. In many programming languages, such as Java, generics are implemented by \emph{erasure}\ie replaced by their upper bound type during compilation into bytecode. This is safe at source level, since the compiler takes care of checking that types are correct, before erasure. However, the types of the generated bytecode are erased and consequently weaker. In a permissionless blockchain, where every user can call the bytecode of smart contracts installed by other users, these weaker types pose a risk of attack. In this paper we start from a real security issue that we found while using generics for writing smart contracts that implement \emph{shared entities} for the Hotmoka blockchain, and we propose a possible solution based on an appropriate code rewriting.

>>>>

This paper shows that generic types (\emph{generics}) are useful for writing more abstract and more general smart contracts, but that this comes with some security risk, reporting a concrete security issue found while using generics for writing smart contracts that implement \emph{shared entities} for the Hotmoka blockchain. That issue can be used to steal the remuneration of validator nodes. This paper proposes a patch based on appropriate code rewriting. Namely, smart contracts are pieces of code that are deployed and executed in the context of a blockchain infrastructure in order to automatically enforce some effects when particular events occur. The writing of smart contracts is a complex and critical activity that can benefit from the use of high-level features of programming languages, and generics is one of them. In many programming languages, such as Java, generics are implemented by \emph{erasure}\ie replaced by their upper bound type during compilation into bytecode. This is safe at source level, since the compiler takes care of checking that types are correct, before erasure. However, the erased types of the generated bytecode are consequently weaker. In a permissionless blockchain, where every user can call the bytecode of smart contracts installed by other users, these weaker types pose a risk of attack.

### 2022-04-27 cluster_computing/010_introduction.tex
Further edits to the Cluster Computing version

In recent years, blockchain technology has rapidly emerged as a powerful tool for supporting the development of many and innovative services and infrastructures. Blockchain-enabled applications are spreading across diverse sectors such as supply chain, business, healthcare, IoT, privacy, and data management \cite{blockchain_applications}. A blockchain is essentially a \textit{distributed ledger}, namely a database replicated across different locations and synchronized by multiple independent participants. Blockchains exploit the redundant, concurrent execution of the same transactions on a decentralized network of many machines, in order to enforce their execution in accordance with a set of predefined rules. Namely, blockchains make it hard, for a single machine, to disrupt the semantics of transactions or their ordering: a misbehaving single machine gets immediately put out of consensus and isolated. 
%
\begin{figure*}[th]
\centering
\includegraphics[width=0.8\linewidth]{figures/solidity_problem}
\caption{Example of possible problem that can occur in Solidity due to the absence of a strong typing mechanism.}
\label{figure.solidity_problem}
\end{figure*}
%
The key innovation introduced by this technology is a mechanism  able to reach an emergent agreement about a global state without the need for a central authority. Morever, another peculiarity is that the consensus is not explicit, because there is not a fixed moment when it occurs.

The rules of blockchain transactions are specified by \emph{smart contracts},
that are code written in a variety of programming languages. To the best of our knowledge, none of them allows generic types and, in any case,
nothing has been published about the opportunity but also the risks of using generic types for writing smart contracts.
The contribution of this paper is exactly to show a real-life
use of generic types for an actual smart contract contained in the support
library of the Takamaka language for writing smart contracts~\cite{Spoto19,Spoto20},
and to demonstrate that a na\"{i}ve use
of Java generics can lead to a code security vulnerability that
allows an attacker to earn money by exploiting someone else's work, with both economical and legal side effects.
This paper provides a fix to that specific issue,
by proposing a re-engineering of the code that forces the compiler to generate defensive checks.
More generally, this paper can be useful for the definition of
bytecode languages for future smart contract languages, by
learning from the weaknesses of Java bytecode, in particualr those related to the
compilation of generics.

@@

The world computer built by Ethereum is known as Ethereum Virtual Machine (EVM) which is the platform where accounts and smart contracts live and are executed.
Solidity is a high-level programming language and smart contracts need to be compiled into bytecode to be executed inside the EVM. At this regard, observe that, in Solidity's bytecode, non-primitive values are referenced through a very general \<address> type. For instance, a Solidity method \<child(Person p, uint256 n) returns Person> actually compiles into \<child(address p, uint256 n) returns address>, losing most type information~\cite{CrafaPZ19}. Since, at run time, it is the bytecode that gets executed, everything can be passed for \<p>, not just a \<Person> instance, as illustrated in Fig.\,\ref{figure.solidity_problem}.
The compiler cannot even enforce strong typing by generating defensive type instance checks and casts, because values are unboxed in Ethereum: they have no attached type information at run time, they are just numerical \emph{addresses}. It follows that, inside the \<child> method, an eventual call to a \<Person>'s method on \<p> might actually execute any arbitrary code, if \<p> is not a \<Person>. In other words, Solidity is not strongly typed. Consequently, it is highly discouraged, in Solidity, to call methods on parameters passed to another method, such as on \<p> passed to \<child>, since an attacker can pass crafted objects for \<p>, with arbitrary implementations for their methods, which can result in the unexpected execution of dangerous code. This actually happened in the case of the infamous DAO hack~\cite{dao16}, that costed millions of dollars.

Strong typing is one of the reasons that pushed towards the adoption
of \emph{traditional} programming languages for smart contracts. For instance,
the Cosmos blockchain~\cite{cosmos} uses Go. The
Hotmoka blockchain~\cite{hotmoka} uses a subset of Java
for smart contracts, called Takamaka~\cite{Spoto19,Spoto20}.
Hyperledger~\cite{hyperldeger} allows Go and Java.
Another reason is the availability of modern
language features, that are missing in Solidity,
such as \emph{generics}\ie the possibility of using
generic type variables. Generics are a powerful and very useful facility for programming
smart contracts, since they allow one to personalize the behavior of such contracts and partially overcome their inherent incompleteness~\cite{ebp}. 
%
Through the use of generics, it is possible to provide to users a set of predefined contract templates that can be extended and specialized by users with lower programming skills, but higher knowledge about the specific application domain. They are based on the use of type placeholders in order
to produce parametrized code, that can be instantiated for each
concrete type provided for the placeholders.
%
However, strong typing and generics are two intertwined language features that have to be carefully considered when smart contracts are implemented and their bytecode is subsequently deployed on a blockchain. For instance, in Java source code, generics are strongly typed, if no \emph{unchecked operations} are used~\cite{NaftalinW06}, as it will always be the case in this paper.
However, generics might have security issues at the level of compiled Java code and this paper originated from a real issue that has been found in our code.

The remainder of this paper is organized as follows.
Sec.~\ref{sec:java_generics} discusses the management of generics in Java.
Sec.~\ref{sec:takamaka} presents the basic notions about the Takamaka language for smart contracts in Java.
Sec.~\ref{sec:shared_entities} shows our real-life Java smart
contracts for shared entities, that use generic types.
Sec.~\ref{sec:validators} shows the instantiation of the shared entities to implement the validators' set
of a proof of stake blockchain.
Sec.~\ref{sec:attack} shows that a na\"{i}ve
deployment of a subclass of the validators' set leads to a code vulnerability.
Sec.~\ref{sec:fix} presents a fix to that vulnerability.
Sec.~\ref{sec:related_work} discusses some related work.
Sec.~\ref{sec:conclusion} concludes.

>>>>

In recent years, blockchain technology has rapidly emerged as a powerful tool for supporting the development of many and innovative services and infrastructures. Blockchain-enabled applications are spreading across diverse sectors such as supply chain, business, healthcare, IoT, privacy, and data management \cite{blockchain_applications}. A blockchain is essentially a \textit{distributed ledger}, namely a database replicated across different locations and synchronized by multiple independent participants. Blockchains exploit the redundant, concurrent execution of the same transactions on a decentralized network of many machines, in order to enforce their execution in accordance with a set of predefined rules. Namely, blockchains make it hard, for a single machine, to disrupt the semantics of transactions or their ordering: a misbehaving single machine gets immediately put out of consensus and isolated. 
%
\begin{figure*}[th]
\centering
\includegraphics[width=0.8\linewidth]{figures/solidity_problem}
\caption{Example of possible problem that can occur in Solidity due to the absence of a strong typing mechanism.}
\label{figure.solidity_problem}
\end{figure*}
%
That is, the key innovation introduced by this technology is a mechanism  able to reach an emergent agreement about a global state without the need for a central authority. Morever, another peculiarity is that the consensus is not explicit, because there is not a fixed moment when it occurs.

The rules of blockchain transactions are specified by \emph{smart contracts},
that are code written in a variety of programming languages. To the best of our knowledge, none of them allows generic types (\emph{generics})
and, in any case,
nothing has been published about the opportunity but also the risks of using generics for writing smart contracts.
The contribution of this paper is exactly to show a real-life
use of generics for an actual smart contract contained in the support
library of the Takamaka language~\cite{Spoto19,Spoto20},
and to demonstrate that a na\"{i}ve use
of Java generics can lead to a code security vulnerability that
allows an attacker to earn money by exploiting someone else's work, with both economical and legal side effects.
This paper provides a fix to that specific issue,
by proposing a re-engineering of the code that forces the compiler to generate defensive checks.
More generally, this paper can be useful for the definition of
future bytecode languages for smart contract languages, by
learning from the weaknesses of Java bytecode, in particular from those related to the
compilation of generics.

@@

The world computer built by Ethereum is known as Ethereum Virtual Machine (EVM) and is the platform where accounts and smart contracts live and are executed.
Solidity is a high-level programming language and smart contracts need to be compiled into bytecode to be executed inside the EVM. At this regard, observe that, in Solidity's bytecode, non-primitive values are referenced through a very general \<address> type. For instance, a Solidity method \<child(Person p, uint256 n) returns Person> actually compiles into \<child(address p, uint256 n) returns address>, losing most type information~\cite{CrafaPZ19}. Since, at run time, it is the bytecode that gets executed, everything can be passed for \<p>, not just a \<Person> instance, as illustrated in Fig.\,\ref{figure.solidity_problem}.
The compiler cannot even enforce strong typing by generating defensive type instance checks and casts, because values are unboxed in Ethereum: they have no attached type information at run time, they are just numerical \emph{addresses}. It follows that, inside the \<child> method, an eventual call to a \<Person>'s method on \<p> might actually execute any arbitrary code, if \<p> is not a \<Person>. In other words, Solidity is not strongly typed. Consequently, it is highly discouraged, in Solidity, to call methods on parameters passed to another method, such as on \<p> passed to \<child>, since an attacker can pass crafted objects for \<p>, with arbitrary implementations for their methods, which can result in the unexpected execution of dangerous code. This actually happened in the case of the infamous DAO hack~\cite{dao16}, that costed millions of dollars.

Strong typing is one of the reasons that pushed towards the adoption
of \emph{traditional} programming languages for smart contracts. For instance,
the Cosmos blockchain~\cite{cosmos} uses Go. The
Hotmoka blockchain~\cite{hotmoka} uses a subset of Java
for smart contracts, called Takamaka~\cite{Spoto19,Spoto20}.
Hyperledger~\cite{hyperldeger} allows Go and Java.
Another reason is the availability of modern
language features, that are missing in Solidity,
such as generics. They are a powerful and very useful facility for programming
smart contracts, since they allow one to personalize the behavior of such contracts and partially overcome their inherent incompleteness~\cite{ebp}. 
%
Through the use of generics, it is possible to provide to users a set of predefined contract templates that they can extend and specialize with lower programming skills, but higher knowledge about the specific application domain. Generics are based on the use of type placeholders in order
to produce parametrized code, that can be instantiated for each
concrete type provided for the placeholders.
%
However, strong typing and generics are two intertwined language features that have to be carefully considered when smart contracts are implemented and their bytecode is subsequently deployed on a blockchain. For instance, in Java source code, generics are strongly typed, if no \emph{unchecked operations} are used~\cite{NaftalinW06}, as it will always be the case in this paper.
However, generics might have security issues at the level of compiled Java code and this paper originated from a real issue that has been found in our code.

The remainder of this paper is organized as follows.
Sec.~\ref{sec:java_generics} discusses the management of generics in Java.
Sec.~\ref{sec:takamaka} presents the basic notions about the Takamaka language for smart contracts in Java.
Sec.~\ref{sec:shared_entities} shows our real-life Java smart
contracts for shared entities, that use generics.
Sec.~\ref{sec:validators} shows the instantiation of the shared entities to implement the validators' set
of a proof of stake blockchain.
Sec.~\ref{sec:attack} shows that a na\"{i}ve
deployment of a subclass of the validators' set leads to a code vulnerability
due to the way generics are compiled.
Sec.~\ref{sec:fix} presents a fix to that vulnerability.
Sec.~\ref{sec:related_work} discusses some related work.
Sec.~\ref{sec:conclusion} concludes.

### 2022-04-27 cluster_computing/020_shared_entities_with_generics.tex
Further edits to the Cluster Computing version

\section{A Generics Shared Entities Implementation}\label{sec:shared_entities}

@@

Fig.~\ref{fig:simple_shared_entity} shows a portion of our \<SimpleSharedEntity>
implementation of the \<SharedEntity> interface in Fig.~\ref{fig:shared_entity}, that
uses two fields: \<shares> maps each shareholder to the amount of shares the it holds and
\<offers> collects the offers that have been placed.
The constructor initially populates the map \<shares> with the initial shareholder.
Other shareholders can be added later, by buying shares\footnote{In the real code, the class has
  constructors to create shared entities with a \emph{set} of initial shareholders. This paper
  reports a simplification of the actual code.}.
Method \<sharesOf> simply accesses
\<shares>, by using zero as default. Method \<place> requires its \<caller()> to be
the seller identified in the \<offer>. This forbids shareholders to sell shares on behalf of others.
Moreover, this guarantees that the caller has type \<S>, the type of \<offer.seller>.
As it has been said before, this cannot be expressed with the syntax of the language.
Method \<place> further requires the seller to be a shareholder with at least \<offer.sharesOnSale>
shares not yet placed on sale. This forbids to oversell more shares
than one owns. At the end, \<place> adds the \<offer> to the set of \<offers>.
Method \<accept> requires that who calls the method must be \<buyer>. Hence, successful
calls to \<accept> can only pass the same caller for \<buyer>. This is a trick to enforce the
caller to have type \<S>, since the syntax of the language does not allow one to express it,
as explained before. Then \<accept> requires the \<offer> to exist, to be still ongoing
and to cost no more than the \<amount> of money provided to \<accept>. If that is the case,
the \<offer> is removed from the \<offers>, shares are moved from seller to buyer (code not
shown in Fig.~\ref{fig:simple_shared_entity}) and the seller of the \<offer>
receives the required price \<offer.cost>.

>>>>

\section{A Generic Shared Entities Implementation}\label{sec:shared_entities}

@@

Fig.~\ref{fig:simple_shared_entity} shows a portion of the code of our \<SimpleSharedEntity>
implementation of the \<SharedEntity> interface in Fig.~\ref{fig:shared_entity}, that
uses two fields: \<shares> maps each shareholder to the amount of shares the it holds and
\<offers> collects the offers that have been placed.
The constructor initially populates the map \<shares> with the initial shareholder.
Other shareholders can be added later, by buying shares\footnote{In the real code, the class has
  constructors to create shared entities with a \emph{set} of initial shareholders. This paper
  reports a simplification of the actual code.}.
Method \<sharesOf> simply accesses
\<shares>, by using zero as default. Method \<place> requires its \<caller()> to be
the seller identified in the \<offer>. This forbids shareholders to sell shares on behalf of others.
Moreover, this guarantees that the caller has type \<S>, the type of \<offer.seller>.
As it has been said before, this cannot be expressed with the syntax of the language.
Method \<place> further requires the seller to be a shareholder with at least \<offer.sharesOnSale>
shares not yet placed on sale. This forbids to oversell more shares
than one owns. At the end, \<place> adds the \<offer> to the set of \<offers>.
Method \<accept> requires that who calls the method must be \<buyer>. Hence, successful
calls to \<accept> can only pass the same caller for \<buyer>. This is a trick to enforce the
caller to have type \<S>, since the syntax of the language does not allow one to express it,
as explained before. Then \<accept> requires the \<offer> to exist, to be still ongoing
and to cost no more than the \<amount> of money provided to \<accept>. If that is the case,
the \<offer> is removed from the \<offers>, shares are moved from seller to buyer (code not
shown in Fig.~\ref{fig:simple_shared_entity}) and the seller of the \<offer>
receives the required price \<offer.cost>.

### 2022-04-27 cluster_computing/030_generics_attack.tex
Further edits to the Cluster Computing version

Let us state an important property about shared entities:
%
\vspace{2ex}
\begin{mdframed}[leftmargin=10pt,rightmargin=10pt]
  \begin{center}\emph{Consistency of Shareholders}\end{center}
  \noindent
  If \<se> is a {\codesize\texttt{SharedEntity<S,O>}}
  then \<se.getShareholders()> contains only elements of type \<S>.
\end{mdframed}
\vspace{2ex}
%
This property is important since it states that one can trust the type \<S> of
the shareholders: if one creates a \<SharedEntity> and fixes a specific type \<S>
for its shareholders, then only instances of \<S> will actually manage to become shareholders.

@@

\begin{figure*}[ht]
\centering
\includegraphics[width=0.9\linewidth]{figures/attack}
\caption{Example of possible attack to a smart contract using Java generics.}
\label{figure.attack}
\end{figure*}

>>>>

Let us state an important, expected property about shared entities:
%
\vspace{2ex}
\begin{mdframed}[leftmargin=10pt,rightmargin=10pt]
  \begin{center}\emph{Consistency of Shareholders}\end{center}
  \noindent
  If \<se> is a {\codesize\texttt{SharedEntity<S,O>} object,}
  then \<se.getShareholders()> contains only elements of type \<S>.
\end{mdframed}
\vspace{2ex}
%
This property is important since it states that one can trust the type \<S> of
the shareholders: if one creates a \<SharedEntity> and fixes a specific type \<S>
for its shareholders, then only instances of \<S> will actually manage to become shareholders.

@@

\begin{figure*}[ht]
\centering
\includegraphics[width=0.9\linewidth]{figures/attack}
\caption{Example of possible attack to a smart contract that uses Java generics.}
\label{figure.attack}
\end{figure*}

### 2022-04-27 cluster_computing/050_related_work.tex
Further edits to the Cluster Computing version

It has been estimated that, on average, software developers make from 100 to 150 errors
for every thousand lines of code~\cite{software_engineering}.
In 2002, the National Institute of Standards and Technology (NIST) estimates that the economic
costs of faulty software in the US is about tens of billions of dollars per year and represent
approximately just under one percent of the Nation's gross domestic product.
The effects induced by errors in software development are even worse when such pieces of software
are smart contracts. Indeed, it is usually impossible to change a smart contract once
it has been deployed, the immutability being one of its main characteristics, so that
errors are treated as intended behaviors. Moreover, smart contracts often store and
manage critical data such as money, digital assets and identities. For this reason, smart contracts
vulnerabilities and correctness are becoming important in literature~\cite{smart_contracts_verification}. Possible solutions can be
classified into three main categories: (i) static analysis of EVM bytecode,
(ii) automatic rectification of EVM bytecode and (iii) development of new languages
for smart contracts.

>>>>

Programming languages specific to smart contracts (such as Solidity)
do not have generic types. General-purpose programming languages do have
generic types in most cases, but are much less frequently used for writing smart contracts.
In any case, we are not aware of any scientific work on the use of generic
types for writing more generic smart contracts, nor of any study on the security
risks, and their solutions, that this implies for the resulting smart contracts. From this point of view,
the present paper has no direct literature to compare with.

It is possible instead to insert this paper in the broader context of software correctness and security.
Namely, it has been estimated that, on average, software developers make from 100 to 150 errors
for every thousand lines of code~\cite{software_engineering}.
In 2002, the National Institute of Standards and Technology (NIST) estimates that the economic
costs of faulty software in the US is about tens of billions of dollars per year and represent
approximately just under one percent of the Nation's gross domestic product.
The effects induced by errors in software development are even worse when such pieces of software
are smart contracts. Indeed, it is usually impossible to change a smart contract once
it has been deployed, the immutability being one of its main characteristics, so that
errors are treated as intended behaviors. Moreover, smart contracts often store and
manage critical data such as money, digital assets and identities. For this reason, smart contracts
vulnerabilities and correctness are becoming important in literature~\cite{smart_contracts_verification}. Possible solutions can be
classified into three main categories: (i) static analysis of EVM bytecode,
(ii) automatic rectification of EVM bytecode and (iii) development of new languages
for smart contracts.

### 2022-04-27 cluster_computing/090_conclusion.tex
Further edits to the Cluster Computing version

This paper has shown that generic types can be useful in the definition
of smart contracts and can simplify the development of rather complex
code such as those for shared entities, and support code reuse, for
instance to implement the validators set of a blockchain network.
However, this paper has shown that generic types
introduce risks of security as well.
Namely, many
programming languages, including Java, erase them at compile time
into types that might be too permissive for low-level calls, such as
those that are started by blockchain transactions. Note that the use of a programming
language without generics is not the solution: Solidity has no generics
and consequently erases \emph{all} reference types into \<address>.
That is the worst possible erasure.

The solution in this paper has been to redefine a method with an argument
of generic type in such a way to call its superclass
(see \<accept> in Fig.~\ref{fig:solution}). This fixes the security risk,
but cannot be regarded as the definite solution to the problem. It is just a trick that
works because it forces the compiler to generate some specific kind of bytecode.
A \emph{smarter} compiler might recognize the redefined \<accept> as
\emph{useless} and just remove it. This would recreate the issue that
has been just solved. That is, the solution in this paper works only for the way
compilers compile \emph{today}.

With hindsight, it is questionable to have implemented generics by erasure
and code instrumentation (bridge methods). If generics would be present
and checked at bytecode level, the attack in Sec.~\ref{sec:attack} would just
be impossible. Currently, generics can only exist as bytecode annotations
that are not mandatory and are ignored by the Java virtual machine
that runs the bytecode. The same consideration might be applied beyond
generics: many features of modern programming
languages have no direct low-level
counterpart but are implemented via instrumentation. Examples are
inner classes and closures (lambda expressions). This is fine at source level, but allows
low-level calls to easily circumvent
the encapsulation guarantees of the language. When embedded in a permissionless
blockchain, such features become dangerous attack surfaces. This paper has shown
the attack surface due to redefinition of methods with a ganeric parameter.
But another example is the use of instrumented methods to allow the access to
private state from inner classes: since inner classes are compiled into distinct
bytecode classes, the compiler adds non-private accessors to the
private state. These accessors cannot be used
at source level, but can be called at bytecode level to gain access to private state.
This paper does not provide a solution to this other issue, but this further example
makes it clear that the attack surface is larger than what described here.

>>>>

This paper has shown that generics are useful in the definition
of smart contracts and can simplify the development of rather complex
code such as that for shared entities, and support code reuse, for
instance to implement the validators set of a blockchain network.
However, this paper has shown that generic types
introduce risks of security as well.
Namely, many
programming languages, including Java, erase them at compile time
into types that might be too permissive for low-level calls, such as
those that are started by blockchain transactions. Note that the use of a programming
language without generics is not the solution: Solidity has no generics
and consequently erases \emph{all} reference types into \<address>.
That is the worst possible erasure.

The solution in this paper has been to redefine the methods that have an argument
of generic type, in such a way to call their superclass
(see the case of \<accept> in Fig.~\ref{fig:solution}). This fixes the security risk,
but cannot be regarded as the definite solution to the problem. It is just a trick that
works because it forces the compiler to generate some specific kind of bytecode.
A \emph{smarter} compiler might recognize the redefined \<accept> as
\emph{useless} and just remove it. This would recreate the issue that
has been just solved. That is, the solution in this paper works only for the way
compilers compile \emph{today}.

With hindsight, it is questionable to have implemented generics by erasure
and code instrumentation (bridge methods). If generics would be present
and checked at bytecode level, the attack in Sec.~\ref{sec:attack} would just
be impossible. Currently, generics can only exist as bytecode annotations
that are not mandatory and are ignored by the Java virtual machine
that runs the bytecode. The same consideration might be applied beyond
generics: many features of modern programming
languages have no direct low-level
counterpart but are implemented via instrumentation. Examples are
inner classes and closures (lambda expressions). This is fine at source level, but allows
low-level calls to easily circumvent
the encapsulation guarantees of the language. When embedded in a permissionless
blockchain, such features become dangerous attack surfaces. This paper has shown
the attack surface due to redefinition of methods with a generic parameter.
But another example is the use of instrumented methods to allow access to
private state from inner classes: since inner classes are compiled into distinct
bytecode classes, the compiler adds non-private accessors to the
private state. These accessors cannot be used
at source level, but can be called at bytecode level to gain access to private state.
This paper does not provide a solution to this other issue, but this further example
makes it clear that the attack surface is larger than what described here.
